---
title: "DSC630 Final Project - Crime Analysis - Part1"
author: "Amie Davis"
date: "27 June, 2020"
output: 
  word_document: default
---
# Data Sources:
Uniform Crime Reporting Program Data: National Incident-Based Reporting System, [United States], 2016;
United States Federal Bureau of Investigation;
Inter-university Consortium for Political and Social Research (ICPSR), University of Michigan;
https://www.icpsr.umich.edu/icpsrweb/NACJD/NIBRS/
\newline\
Geodetic Data for US Cities:
https://simplemaps.com/data/us-cities

# Load Libraries

```{r libs, message=FALSE}
library(readr)
library(ggplot2)
library(lubridate)
library(dplyr)
require(scales)
```

# 1. Prepare Data

## a) Import the Data
```{r load, message=FALSE}

# Load victim data
vic_data <- read_tsv("Data/UCR_2016/Victim_Segment/37065-0005-Data.tsv",
  col_types = cols(
  .default = col_character(),
  V4024 = col_double(),
  V4010 = col_character()
) )

# Load precinct data
bat_data <- read_tsv("Data/UCR_2016/Batch_Header_Segment/37065-0001-Data.tsv")
```

## b) Review Features

```{r vic_str, eval=FALSE}
str(vic_data)
```

```{r bat_str, eval=FALSE}
str(bat_data)
```

## c) Derived Features

```{r vic_derived}
# Convert date fields to date stamps
vic_date_data <- vic_data %>% mutate(V4005 = ymd(V4005))

# Split date fields into separate columns using lubridate package
vic_new_data <- vic_date_data %>% mutate (VIC_INC_YEAR = year(V4005), 
                                VIC_INC_MONTH = month(V4005), 
                                VIC_INC_DAY = day(V4005),
                                VIC_INC_DOW = weekdays(V4005))

# Convert categorical variables to factors
vic_new_data$VIC_INC_MONTH <- factor(vic_new_data$VIC_INC_MONTH)
vic_new_data$VIC_INC_DOW <- factor(vic_new_data$VIC_INC_DOW)

vic_new_data$V4007 <- factor(vic_new_data$V4007)
vic_new_data$V4008 <- factor(vic_new_data$V4008)
vic_new_data$V4009 <- factor(vic_new_data$V4009)
vic_new_data$V4010 <- factor(vic_new_data$V4010)
vic_new_data$V4011 <- factor(vic_new_data$V4011)
vic_new_data$V4012 <- factor(vic_new_data$V4012)
vic_new_data$V4013 <- factor(vic_new_data$V4013)
vic_new_data$V4014 <- factor(vic_new_data$V4014)
vic_new_data$V4015 <- factor(vic_new_data$V4015)
vic_new_data$V4016 <- factor(vic_new_data$V4016)
vic_new_data$V4017 <- factor(vic_new_data$V4017)
vic_new_data$V4017A <- factor(vic_new_data$V4017A)
vic_new_data$V4017B <- factor(vic_new_data$V4017B)
vic_new_data$V4019 <- factor(vic_new_data$V4019)
vic_new_data$V4020 <- factor(vic_new_data$V4020)
vic_new_data$V4021 <- factor(vic_new_data$V4021)
vic_new_data$V4022 <- factor(vic_new_data$V4022)
vic_new_data$V4023 <- factor(vic_new_data$V4023)
vic_new_data$V4024 <- factor(vic_new_data$V4024)
vic_new_data$V4025 <- factor(vic_new_data$V4025)
vic_new_data$V4026 <- factor(vic_new_data$V4026)
vic_new_data$V4027 <- factor(vic_new_data$V4027)
vic_new_data$V4028 <- factor(vic_new_data$V4028)
vic_new_data$V4029 <- factor(vic_new_data$V4029)
vic_new_data$V4030 <- factor(vic_new_data$V4030)
vic_new_data$V4032 <- factor(vic_new_data$V4032)
vic_new_data$V4034 <- factor(vic_new_data$V4034)
vic_new_data$V4036 <- factor(vic_new_data$V4036)
vic_new_data$V4038 <- factor(vic_new_data$V4038)
vic_new_data$V4040 <- factor(vic_new_data$V4040)
vic_new_data$V4042 <- factor(vic_new_data$V4042)
vic_new_data$V4044 <- factor(vic_new_data$V4044)
vic_new_data$V4046 <- factor(vic_new_data$V4046)
vic_new_data$V4048 <- factor(vic_new_data$V4048)
vic_new_data$V4050 <- factor(vic_new_data$V4050)

#head(vic_new_data)
```

```{r bat_derived}

# Convert categorical variables to factors

# Location Groupings
bat_data$BH002 <- factor(bat_data$BH002)
bat_data$BH008 <- factor(bat_data$BH008)
bat_data$BH009 <- factor(bat_data$BH009)
bat_data$BH010 <- factor(bat_data$BH010)
bat_data$BH011 <- factor(bat_data$BH011)
bat_data$BH012 <- factor(bat_data$BH012)
bat_data$BH013 <- factor(bat_data$BH013)

# Fed District
bat_data$BH015 <- factor(bat_data$BH015)
bat_data$BH016 <- factor(bat_data$BH016)

# MSA and Country Codes
bat_data$BH020 <- factor(bat_data$BH020)
bat_data$BH021 <- factor(bat_data$BH021)
bat_data$BH023 <- factor(bat_data$BH023)
bat_data$BH025 <- factor(bat_data$BH025)
bat_data$BH028 <- factor(bat_data$BH028)
bat_data$BH029 <- factor(bat_data$BH029)
bat_data$BH032 <- factor(bat_data$BH032)
bat_data$BH033 <- factor(bat_data$BH033)
bat_data$BH036 <- factor(bat_data$BH036)
bat_data$BH037 <- factor(bat_data$BH037)

# FIPS County Codes
bat_data$BH054 <- factor(bat_data$BH054)
bat_data$BH055 <- factor(bat_data$BH055)
bat_data$BH056 <- factor(bat_data$BH056)
bat_data$BH057 <- factor(bat_data$BH057)
bat_data$BH058 <- factor(bat_data$BH058)

#head(bat_data)
```

## d) Re-Label data fields

```{r new_vic_vars, include=TRUE}
names(vic_new_data)[names(vic_new_data) == "V4001"] <- "ARR_SEG_LEVEL"   #NOT NEEDED - ALWAYS 4 FOR VICTIM FILE
names(vic_new_data)[names(vic_new_data) == "V4002"] <- "VIC_STATE_CODE"
names(vic_new_data)[names(vic_new_data) == "V4003"] <- "ORI"
names(vic_new_data)[names(vic_new_data) == "V4004"] <- "INC_NUM" 
names(vic_new_data)[names(vic_new_data) == "V4005"] <- "VIC_INC_DATE" #DATE
names(vic_new_data)[names(vic_new_data) == "V4006"] <- "VIC_SEQ_NUM"

names(vic_new_data)[names(vic_new_data) == "V4007"] <- "OFF_CODE01"
names(vic_new_data)[names(vic_new_data) == "V4008"] <- "OFF_CODE02"
names(vic_new_data)[names(vic_new_data) == "V4009"] <- "OFF_CODE03"
names(vic_new_data)[names(vic_new_data) == "V4010"] <- "OFF_CODE04"
names(vic_new_data)[names(vic_new_data) == "V4011"] <- "OFF_CODE05"
names(vic_new_data)[names(vic_new_data) == "V4012"] <- "OFF_CODE06"
names(vic_new_data)[names(vic_new_data) == "V4013"] <- "OFF_CODE07"
names(vic_new_data)[names(vic_new_data) == "V4014"] <- "OFF_CODE08"
names(vic_new_data)[names(vic_new_data) == "V4015"] <- "OFF_CODE09"
names(vic_new_data)[names(vic_new_data) == "V4016"] <- "OFF_CODE10"
names(vic_new_data)[names(vic_new_data) == "V4017"] <- "VICTIM_TYPE"        #CAT - 9
names(vic_new_data)[names(vic_new_data) == "V4017A"] <- "ACT_TYPE_OFFC"     #CAT - 11 (numeric)
names(vic_new_data)[names(vic_new_data) == "V4017B"] <- "ASSG_TYPE_OFFC"    #CAT - 7
names(vic_new_data)[names(vic_new_data) == "V4017C"] <- "ORI_OTHER" 

names(vic_new_data)[names(vic_new_data) == "V4018"] <- "AGE_OF_VICTIM"      #NUM (00-UNK, 99-99+)
names(vic_new_data)[names(vic_new_data) == "V4019"] <- "SEX_OF_VICTIM"      #CAT (M/F)
names(vic_new_data)[names(vic_new_data) == "V4020"] <- "RACE_OF_VICTIM"     #CAT - 5 VALUES, U=UNK
names(vic_new_data)[names(vic_new_data) == "V4021"] <- "ETHNIC_OF_VIC"      #CAT (H/N/U)
names(vic_new_data)[names(vic_new_data) == "V4022"] <- "VIC_RESIDENT"       #CAT (R/N/U)

names(vic_new_data)[names(vic_new_data) == "V4023"] <- "ASSAULT_CIRC1"      #CAT - 17
names(vic_new_data)[names(vic_new_data) == "V4024"] <- "ASSAULT_CIRC2"      #CAT - 17
names(vic_new_data)[names(vic_new_data) == "V4025"] <- "JUST_HOM_CIRC"      #CAT - 7

names(vic_new_data)[names(vic_new_data) == "V4026"] <- "INJURY_TYPE1"       #CAT - 8
names(vic_new_data)[names(vic_new_data) == "V4027"] <- "INJURY_TYPE2"       #CAT - 8
names(vic_new_data)[names(vic_new_data) == "V4028"] <- "INJURY_TYPE3"       #CAT - 8
names(vic_new_data)[names(vic_new_data) == "V4029"] <- "INJURY_TYPE4"       #CAT - 8
names(vic_new_data)[names(vic_new_data) == "V4030"] <- "INJURY_TYPE5"       #CAT - 8

names(vic_new_data)[names(vic_new_data) == "V4031"] <- "OFF_NUM_KEY1" 
names(vic_new_data)[names(vic_new_data) == "V4032"] <- "REL_TO_OFF1"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4033"] <- "OFF_NUM_KEY2" 
names(vic_new_data)[names(vic_new_data) == "V4034"] <- "REL_TO_OFF2"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4035"] <- "OFF_NUM_KEY3"
names(vic_new_data)[names(vic_new_data) == "V4036"] <- "REL_TO_OFF3"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4037"] <- "OFF_NUM_KEY4" 
names(vic_new_data)[names(vic_new_data) == "V4038"] <- "REL_TO_OFF4"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4039"] <- "OFF_NUM_KEY5" 
names(vic_new_data)[names(vic_new_data) == "V4040"] <- "REL_TO_OFF5"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4041"] <- "OFF_NUM_KEY6" 
names(vic_new_data)[names(vic_new_data) == "V4042"] <- "REL_TO_OFF6"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4043"] <- "OFF_NUM_KEY7" 
names(vic_new_data)[names(vic_new_data) == "V4044"] <- "REL_TO_OFF7"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4045"] <- "OFF_NUM_KEY8" 
names(vic_new_data)[names(vic_new_data) == "V4046"] <- "REL_TO_OFF8"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4047"] <- "OFF_NUM_KEY9" 
names(vic_new_data)[names(vic_new_data) == "V4048"] <- "REL_TO_OFF9"        #CAT - 26
names(vic_new_data)[names(vic_new_data) == "V4049"] <- "OFF_NUM_KEY10"
names(vic_new_data)[names(vic_new_data) == "V4050"] <- "REL_TO_OFF10"       #CAT - 26

names(vic_new_data)[names(vic_new_data) == "V4051"] <- "NUM_RECS_PER_VICTIM" 

#str(vic_new_data)
```

```{r bat_vars, include=TRUE}
names(bat_data)[names(bat_data) == "BH002"] <- "NUM_STATE_CODE"    # CAT (numeric)
names(bat_data)[names(bat_data) == "BH006"] <- "DATE_WENT_NBIRS"   # DATE
names(bat_data)[names(bat_data) == "BH003"] <- "ORI"
names(bat_data)[names(bat_data) == "BH007"] <- "CITY_NAME" 
names(bat_data)[names(bat_data) == "BH008"] <- "STATE_ABBR"        #CAT
names(bat_data)[names(bat_data) == "BH009"] <- "POP_GROUP"         #CAT - 23
names(bat_data)[names(bat_data) == "BH010"] <- "CTRY_DIVISION"     #CAT - 10
names(bat_data)[names(bat_data) == "BH011"] <- "CTRY_REGION"       #CAT - 5
names(bat_data)[names(bat_data) == "BH012"] <- "AGENCY_IND"        #CAT - 9 (numeric)
names(bat_data)[names(bat_data) == "BH013"] <- "CORE_CITY"         #CAT (Y/N)
names(bat_data)[names(bat_data) == "BH015"] <- "FBI_OFFICE" 
names(bat_data)[names(bat_data) == "BH016"] <- "JUDICIAL_DIST" 
names(bat_data)[names(bat_data) == "BH019"] <- "CURRENT_POP1" 
names(bat_data)[names(bat_data) == "BH020"] <- "UCR_COUNTY_CD1"
names(bat_data)[names(bat_data) == "BH021"] <- "MSA_CD1"
names(bat_data)[names(bat_data) == "BH022"] <- "LAST_POP1"
names(bat_data)[names(bat_data) == "BH023"] <- "CURRENT_POP2" 
names(bat_data)[names(bat_data) == "BH023"] <- "UCR_COUNTY_CD2"
names(bat_data)[names(bat_data) == "BH025"] <- "MSA_CD2"
names(bat_data)[names(bat_data) == "BH026"] <- "LAST_POP2"
names(bat_data)[names(bat_data) == "BH027"] <- "CURRENT_POP3" 
names(bat_data)[names(bat_data) == "BH028"] <- "UCR_COUNTY_CD3"
names(bat_data)[names(bat_data) == "BH029"] <- "MSA_CD3"
names(bat_data)[names(bat_data) == "BH030"] <- "LAST_POP3"
names(bat_data)[names(bat_data) == "BH031"] <- "CURRENT_POP4" 
names(bat_data)[names(bat_data) == "BH032"] <- "UCR_COUNTY_CD4"
names(bat_data)[names(bat_data) == "BH033"] <- "MSA_CD4"
names(bat_data)[names(bat_data) == "BH034"] <- "LAST_POP4"
names(bat_data)[names(bat_data) == "BH035"] <- "CURRENT_POP5" 
names(bat_data)[names(bat_data) == "BH036"] <- "UCR_COUNTY_CD5"
names(bat_data)[names(bat_data) == "BH037"] <- "MSA_CD5"
names(bat_data)[names(bat_data) == "BH038"] <- "LAST_POP5"
names(bat_data)[names(bat_data) == "BH054"] <- "FIPS_COUNTY1"
names(bat_data)[names(bat_data) == "BH055"] <- "FIPS_COUNTY2"
names(bat_data)[names(bat_data) == "BH056"] <- "FIPS_COUNTY3"
names(bat_data)[names(bat_data) == "BH057"] <- "FIPS_COUNTY4"
names(bat_data)[names(bat_data) == "BH058"] <- "FIPS_COUNTY5"

#str(bat_data)
```

## e) Join Datasets
```{r join, include=TRUE}

# Join victim data with ORI reference data
comb_df <- merge(vic_new_data, bat_data, by="ORI")

#head(comb_df)
#summary(comb_df)
```

## f) Drop unneeded columns

```{r drop, include=TRUE}

# After analysis of the codebook, the following fields are not needed and will be removed.
# Dropping file id number, since it is a constant.
# Using state code from batch file vice victim file
# Only using location information from batch file
# Relationship victim has to offender is out of scope of this project, so the related fields will be removed.
# Each reporting district can associate up to 5 metropolitan areas.  For the scope of this project, I will limit to the 1st area.

  comb_df[ ,c(
    "ARR_SEG_LEVEL",
    "VIC_SEQ_NUM",
    "VIC_STATE_CODE",
    "ORI_OTHER",
    "OFF_NUM_KEY1",
    "OFF_NUM_KEY2",
    "OFF_NUM_KEY3",
    "OFF_NUM_KEY4",
    "OFF_NUM_KEY5",
    "OFF_NUM_KEY6",
    "OFF_NUM_KEY7",
    "OFF_NUM_KEY8",
    "OFF_NUM_KEY9",
    "OFF_NUM_KEY10",
    "REL_TO_OFF1",
    "REL_TO_OFF2",
    "REL_TO_OFF3",
    "REL_TO_OFF4",
    "REL_TO_OFF5",
    "REL_TO_OFF6",
    "REL_TO_OFF7",
    "REL_TO_OFF8",
    "REL_TO_OFF9",
    "REL_TO_OFF10",
    "BH001",
    "BH004",
    "BH005",
    "BH014",
    "BH017",
    "BH018",
    "BH024",
    "BH039",
    "BH040",
    "BH041",
    "BH042",
    "BH043",
    "BH044",
    "BH045",
    "BH046",
    "BH047",
    "BH048",
    "BH049",
    "BH050",
    "BH051",
    "BH052",
    "BH053",
    "BH059",
    "BH060",
    "DATE_WENT_NBIRS",
    "CURRENT_POP2", 
    "UCR_COUNTY_CD2",
    "MSA_CD2",
    "LAST_POP2",
    "CURRENT_POP3", 
    "UCR_COUNTY_CD3",
    "MSA_CD3",
    "LAST_POP3",
    "CURRENT_POP4", 
    "UCR_COUNTY_CD4",
    "MSA_CD4",
    "LAST_POP4",
    "CURRENT_POP5", 
    "UCR_COUNTY_CD5",
    "MSA_CD5",
    "LAST_POP5",
    "FIPS_COUNTY2",
    "FIPS_COUNTY3",
    "FIPS_COUNTY4",
    "FIPS_COUNTY5"
)] <- list(NULL)
```

## g) Summary Statistics
```{r summ, eval=FALSE}
summary(comb_df)
```
\newline\
Observations: \newline\
6034725 Victim Records \newline\
Some fields have nothing but N/A values \newline\
Outliers in Victim Date Field \newline\

```{r drop2, include=TRUE}

# After analysis of the summary statistics, the following fields are not needed and will be removed.
# Removing fields without data - only N/A values exist

  comb_df[ ,c(
    "OFF_CODE08",
    "OFF_CODE09",
    "OFF_CODE10",
    "CURRENT_POP4",
    "UCR_COUNTY_CD4",
    "MSA_CD4",
    "LAST_POP4",
    "CURRENT_POP5",
    "UCR_COUNTY_CD5",
    "MSA_CD5",
    "LAST_POP5",
    "FIPS_COUNTY4",
    "FIPS_COUNTY5"
)] <- list(NULL)
  
#summary(comb_df)
```

## h) Remove Outliers
```{r outliers, include=TRUE}

# Remove outliers with incident dates from 2015.  Look at 2016 incidents only.
clean_data <- subset(comb_df, VIC_INC_YEAR == 2016)

summary(clean_data$VIC_INC_DATE)
#str(clean_data)
#5951120 records
```

## i) Handle NA Values in Numeric Fields
```{r hand_na, warning=FALSE}

# 1) Victim's Age
clean_data$AGE_OF_VICTIM <- as.numeric(clean_data$AGE_OF_VICTIM)
summary(clean_data$AGE_OF_VICTIM)

# Distribution is slightly skewed towards mean, so will impute with median
# Impute NAs with median age
clean_data$AGE_OF_VICTIM[is.na(clean_data$AGE_OF_VICTIM)] <- median(clean_data$AGE_OF_VICTIM, na.rm = TRUE)

print('Imputed Summary')
summary(clean_data$AGE_OF_VICTIM)

# 2) Population Fields
summary(comb_df$CURRENT_POP1)
summary(comb_df$LAST_POP1)

# Replace 0 with mean population
clean_data$CURRENT_POP1[clean_data$CURRENT_POP1 == 0] <- NA
clean_data$LAST_POP1[clean_data$LAST_POP1 == 0] <- NA

# Impute with mean population
clean_data$CURRENT_POP1[is.na(clean_data$CURRENT_POP1)] <- mean(clean_data$CURRENT_POP1, na.rm = TRUE)
clean_data$LAST_POP1[is.na(clean_data$LAST_POP1)] <- mean(clean_data$LAST_POP1, na.rm = TRUE)

print('Imputed Summary')
summary(clean_data$CURRENT_POP1)
summary(clean_data$LAST_POP1)

```

## j) Apply filters
```{r jh_filter, include=TRUE}

# Create subset to include only records marked as justifiable homicide
just_hom_df <- filter(clean_data, OFF_CODE01 == "09C" | OFF_CODE02 == "09C" | OFF_CODE03 == "09C" |
                   OFF_CODE04 == "09C" | OFF_CODE05 == "09C" | OFF_CODE06 == "09C" | OFF_CODE07 == "09C")

#summary(just_hom_df)
# There are 217 incidents of justifiable homicide reported to UCR in 2016
```

```{r agg_filter, include=TRUE}

# Create subset to include only records marked as aggravated assault
agg_asst_df <- filter(clean_data, OFF_CODE01 == "13A" | OFF_CODE02 == "13A" | OFF_CODE03 == "13A" |
                   OFF_CODE04 == "13A" | OFF_CODE05 == "13A" | OFF_CODE06 == "13A" | OFF_CODE07 == "13A")

#summary(agg_asst_df)
# There are 235,811 incidents of aggravated assaults reported to UCR in 2016
```

# 2. EDA - Review Distributions

## a) Plot Histograms for Numeric Vars

```{r age_hist}

ggplot(clean_data, aes(x=AGE_OF_VICTIM)) +
    geom_histogram(bins=20) +
    labs(x="Age") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
\newline\
The peak is inflated to to imputation of na values.  Distribution is fairly normal, but is slightly left-skewed.
\newline\

```{r pop_hist}

ggplot(clean_data, aes(x=CURRENT_POP1)) +
    geom_histogram(bins=20) +
    labs(x="Population") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
\newline\
Distribution is left-skewed.
\newline\

## b) Box Plots for Numeric Features

```{r box1}
boxplot(clean_data$AGE_OF_VICTIM,
  main="Victim Age")

boxplot(clean_data$CURRENT_POP1,
  main="District Population")

boxplot(clean_data$LAST_POP1,
  main="Last Census Population")
```

## c) Histograms for Categorical Features
```{r hist_plot1}

# Victim Type
p <- ggplot(clean_data, aes(x=VICTIM_TYPE)) +
     geom_bar(fill="dark green") +
     labs(x="Victim Type", y="Victim Records", title="Records By Victim Type") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
I = Individual \newline\
B = Business \newline\
S = Society/Public
\newline

```{r hist_plot2}

# Activity Type
# Most values are NA.  Removed NA to review remaining fields
p <- ggplot(data=subset(clean_data, !is.na(ACT_TYPE_OFFC)), aes(x=ACT_TYPE_OFFC)) +
     geom_bar(fill="dark green") +
     labs(x="Activity Type", y="Victim Records", title="Records By Activity Type") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
This field is only used when an officer is assaulted or killed in the line of duty \newline\
1 = Respond disturbance call (Family quarrels, person with firearm, etc.) \newline\
2 = Burglaries in progress or pursuing burglary suspects \newline\
3 = Robberies in progress or pursuing robbery suspects \newline\
4 = Attempting other arrests \newline\
5 = Civil disorder (Riot, mass disobedience) \newline\
6 = Handling, transporting, custody of prisoners \newline\
7 = Investigating suspicious persons or circumstances \newline\
8 = Ambush-no warning \newline\
9 = Mentally deranged assailant \newline\
10 = Traffic pursuits and stops \newline\
11 = All other
\newline\

```{r hist_plot3}

# Assignment Type

# Most values are NA.  Removed NA to review remaining fields
p <- ggplot(data=subset(clean_data, !is.na(ASSG_TYPE_OFFC)), aes(x=ASSG_TYPE_OFFC)) +
     geom_bar(fill="dark green") +
     labs(x="Assignment Type", y="Victim Records", title="Records By Assignment Type") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
F = Two-officer vehicle \newline\
G = One-officer vehicle (alone) \newline\
H = One-officer vehicle (assisted) \newline\
I = Detective or special assignment (alone) \newline\
J = Detective or special assignment (assisted) \newline\
K = Other (alone) \newline\
L = Other (assisted) \newline\
\newline\

```{r hist_plot4}

# Sex of Victim
p <- ggplot(clean_data, aes(x=SEX_OF_VICTIM)) +
     geom_bar(fill="dark green") +
     labs(x="Sex of Victim", y="Victim Records", title="Records By Sex of Victim") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```

```{r hist_plot5}

# Race of Victim
p <- ggplot(clean_data, aes(x=RACE_OF_VICTIM)) +
     geom_bar(fill="dark green") +
     labs(x="Victim Race", y="Victim Records", title="Records By Victim Race") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
W = White \newline\
B = Black or African American \newline\
I = American Indian or Alaska Native \newline\
A = Asian \newline\
P = Native Hawaiian or Other Pacific Islander \newline\
U = Unknown \newline\
\newline\

```{r hist_plot6}
p <- ggplot(clean_data, aes(x=ETHNIC_OF_VIC)) +
     geom_bar(fill="dark green") +
     labs(x="Victim Ethnicity", y="Victim Records", title="Records By Victim Ethnicity") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
H = Hispanic or Latino Origin \newline\
N = Not of Hispanic or Latino Origin \newline\
U = Unknown \newline\
\newline\

```{r hist_plot7}
p <- ggplot(clean_data, aes(x=VIC_RESIDENT)) +
     geom_bar(fill="dark green") +
     labs(x="Victim Residence", y="Victim Records", title="Records By Victim Residence") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
R = Resident.  Victim is a resident of the reporting precinct. \newline\
N = Nonresident \newline\
U = Unknown \newline\
\newline\

```{r hist_plot8}

# Justifiable Homicide

# Most values are NA.  Removed NA to review remaining fields
p <- ggplot(data=subset(clean_data, !is.na(JUST_HOM_CIRC)), aes(x=JUST_HOM_CIRC)) +
     geom_bar(fill="dark green") +
     labs(x="Justifiable Homicide Reason", y="Victim Records", title="Records By Justifiable Homicide Reason") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
This field is only used if assault circumstance is justifiable homicide) \newline\
A=Criminal Attacked Police Officer and That Officer Killed Criminal \newline\
B=Criminal Attacked Fellow Police Officer and Criminal Killed by Another Police Officer \newline\
C=Criminal Attacked a Civilian \newline\
D=Criminal Attempted Flight From a Crime \newline\
E=Criminal Killed In Commission of a Crime \newline\
F=Criminal Resisted Arrest \newline\
G=Unable to Determine/Not Enough Information \newline\
\newline\

```{r hist_plot9}

# State

p <- ggplot(clean_data, aes(x=STATE_ABBR)) +
     geom_bar(fill="dark green") +
     labs(x="State", y="Victim Records", title="Records By State") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
Note that some states may have switched over to the NBIRS system.  I may want to focus on states with the highest reported incidents.
\newline\

```{r hist_plot10}

# Population Bin

p <- ggplot(clean_data, aes(x=POP_GROUP)) +
     geom_bar(fill="dark green") +
     labs(x="Population Bin", y="Victim Records", title="Records By Population Bin") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
1B=Cities from 500,000 thru 999,999 \newline\
1C=Cities from 250,000 thru 499,999 \newline\
2=Cities from 100,000 thru 249,999 \newline\
3=Cities from 50,000 thru 99,999 \newline\
4=Cities from 25,000 thru 49,999 \newline\
5=Cities from 10,000 thru 24,999 \newline\
6=Cities from 2,500 thru 9,999 \newline\
7=Cities under 2,500 \newline\
8A=Non-MSA Counties 100,000 or over \newline\
8B=Non-MSA Counties from 25,000 thru 99,999 \newline\
8C=Non-MSA Counties from 10,000 thru 24,999 \newline\
8D=Non-MSA Counties under 10,000 \newline\
8E=Non-MSA State Police \newline\
9A=MSA Counties 100,000 or over \newline\
9B=MSA Counties from 25,000 thru 99,999 \newline\
9C=MSA Counties from 10,000 thru 24,999 \newline\
9D=MSA Counties under 10,000 \newline\
9E=MSA State Police \newline\
\newline\

```{r hist_plot11}

# Country Division

p <- ggplot(clean_data, aes(x=CTRY_DIVISION)) +
     geom_bar(fill="dark green") +
     labs(x="Country Division", y="Victim Records", title="Records By Country Division") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
1=New England \newline\
2=Middle Atlantic \newline\
3=East North Central \newline\
4=West North Central \newline\
5=South Atlantic \newline\
6=East South Central \newline\
7=West South Central \newline\
8=Mountain \newline\
9=Pacific \newline\
\newline\

```{r hist_plot12}

# Country Region

p <- ggplot(clean_data, aes(x=CTRY_REGION)) +
     geom_bar(fill="dark green") +
     labs(x="Country Region", y="Victim Records", title="Records By Country Region") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
1=North East \newline\
2=North Central \newline\
3=South \newline\
4=West \newline\
\newline\

```{r hist_plot13}

# Agency Type

p <- ggplot(clean_data, aes(x=AGENCY_IND)) +
     geom_bar(fill="dark green") +
     labs(x="Agency Type", y="Victim Records", title="Records By Agency Type") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
0=Covered-By Another Agency \newline\
1=City \newline\
2=County \newline\
3=University or College \newline\
4=State Police \newline\
5=Special Agency \newline\
6=Other State Agencies \newline\
7=Tribal Agencies \newline\
\newline\

```{r hist_plot14}

# Core City Indicator

p <- ggplot(clean_data, aes(x=CORE_CITY)) +
     geom_bar(fill="dark green") +
     labs(x="Core City Indicator", y="Victim Records", title="Records By Core City") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```

```{r hist_plot15}

# FBI Office

p <- ggplot(clean_data, aes(x=FBI_OFFICE)) +
     geom_bar(fill="dark green") +
     labs(x="FBI Office", y="Victim Records", title="Records By FBI Office") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```

```{r hist_plot16}

# Judicial District

p <- ggplot(clean_data, aes(x=JUDICIAL_DIST)) +
     geom_bar(fill="dark green") +
     labs(x="Judicial District", y="Victim Records", title="Records By Judicial District") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```

```{r hist_plot17}

# County Code

# Most values are NA.  Removed NA to review remaining fields
p <- ggplot(data=subset(clean_data, !is.na(UCR_COUNTY_CD1)), aes(x=UCR_COUNTY_CD1)) +
     geom_bar(fill="dark green") +
     labs(x="County Code", y="Victim Records", title="Records By County Code") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```

```{r hist_plot18}

# MSA Code

# Most values are NA.  Removed NA to review remaining fields
p <- ggplot(data=subset(clean_data, !is.na(MSA_CD1)), aes(x=MSA_CD1)) +
     geom_bar(fill="dark green") +
     labs(x="MSA Code", y="Victim Records", title="Records By MSA Code") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```

```{r offense, include=TRUE}

# Create new df to include all offenses and counts
off1_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$OFF_CODE01))
off2_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$OFF_CODE02))
off3_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$OFF_CODE03))
off4_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$OFF_CODE04))
off5_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$OFF_CODE05))
off6_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$OFF_CODE06))
off7_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$OFF_CODE07))

names(off1_df)[names(off1_df) == "clean_data.OFF_CODE01"] <- "OFF_CODE"
names(off2_df)[names(off2_df) == "clean_data.OFF_CODE02"] <- "OFF_CODE"
names(off3_df)[names(off3_df) == "clean_data.OFF_CODE03"] <- "OFF_CODE"
names(off4_df)[names(off4_df) == "clean_data.OFF_CODE04"] <- "OFF_CODE"
names(off5_df)[names(off5_df) == "clean_data.OFF_CODE05"] <- "OFF_CODE"
names(off6_df)[names(off6_df) == "clean_data.OFF_CODE06"] <- "OFF_CODE"
names(off7_df)[names(off7_df) == "clean_data.OFF_CODE07"] <- "OFF_CODE"

off_df <- rbind(off1_df, off2_df, off3_df, off4_df, off5_df, off6_df, off7_df) 


#head(off_df)
#summary(off_df)
```

```{r hist_plot19}

# Offense Codes

p <- ggplot(off_df, aes(x=OFF_CODE)) +
     geom_bar(fill="dark green") +
     labs(x="Offense Codes", y="Victim Records", title="Records By Offense Codes") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```

```{r assault, include=TRUE}

# Create new df to include all Aggravated Assault/Homicide Circumstances and counts
asst1_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$ASSAULT_CIRC1))
asst2_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$ASSAULT_CIRC2))

names(asst1_df)[names(asst1_df) == "clean_data.ASSAULT_CIRC1"] <- "ASSAULT_CIRC"
names(asst2_df)[names(asst2_df) == "clean_data.ASSAULT_CIRC2"] <- "ASSAULT_CIRC"

asst_df <- rbind(asst1_df, asst2_df) 


#head(asst_df)
```

```{r hist_plot20}

# Aggravated Assault/Homicide Circumstances

p <- ggplot(asst_df, aes(x=ASSAULT_CIRC)) +
     geom_bar(fill="dark green") +
     labs(x="Circumstance", y="Victim Records", title="Records By Aggravated Assault/Homicide Circumstances") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
01=Argument \newline\
02=Assault on Law Enforcement Officer(s) \newline\
03=Drug Dealing \newline\
04=Gangland \newline\
05=Juvenile Gang \newline\
06=Lovers' Quarrel \newline\
07=Mercy Killing (Not applicable to Aggravated Assault) \newline\
08=Other Felony Involved \newline\
09=Other Circumstances \newline\
10=Unknown Circumstances \newline\
30=Child Playing With Weapon \newline\
31=Gun-Cleaning Accident \newline\
32=Hunting Accident \newline\
33=Other Negligent Weapon Handling \newline\
34=Other Negligent Killings \newline\
20=Criminal Killed by Private Citizen (JUSTIFIABLE HOMICIDE) \newline\
21=Criminal Killed by Police Officer (JUSTIFIABLE HOMICIDE)
\newline\

```{r injury, include=TRUE}

# Create new df to include all offenses and counts
inj1_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$INJURY_TYPE1))
inj2_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$INJURY_TYPE2))
inj3_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$INJURY_TYPE3))
inj4_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$INJURY_TYPE4))
inj5_df <- na.omit(data.frame(clean_data$ORI, clean_data$INC_NUM, clean_data$INJURY_TYPE5))

names(inj1_df)[names(inj1_df) == "clean_data.INJURY_TYPE1"] <- "INJURY_TYPE"
names(inj2_df)[names(inj2_df) == "clean_data.INJURY_TYPE2"] <- "INJURY_TYPE"
names(inj3_df)[names(inj3_df) == "clean_data.INJURY_TYPE3"] <- "INJURY_TYPE"
names(inj4_df)[names(inj4_df) == "clean_data.INJURY_TYPE4"] <- "INJURY_TYPE"
names(inj5_df)[names(inj5_df) == "clean_data.INJURY_TYPE5"] <- "INJURY_TYPE"

inj_df <- rbind(inj1_df, inj2_df, inj3_df, inj4_df, inj5_df) 


#head(off_df)
```
```{r hist_plot21}

# Injury Types

p <- ggplot(inj_df, aes(x=INJURY_TYPE)) +
     geom_bar(fill="dark green") +
     labs(x="Injury Type", y="Victim Records", title="Records By Injury Type") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
N=None \newline\
M=Apparent Minor Injury \newline\
B=Apparent Broken Bones \newline\
O=Other Major Injury \newline\
I=Possible Internal Injury \newline\
T=Loss of Teeth \newline\
L=Severe Laceration \newline\
U=Unconsciousness \newline\
\newline\

```{r hist_plot22}

# Month

p <- ggplot(clean_data, aes(x=VIC_INC_MONTH)) +
     geom_bar(fill="dark green") +
     labs(x="Month", y="Victim Records", title="Records By Month") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
This is surprising.  I would think crime might be lower in winter months and higher in summer.  Although that is slightly true, the difference does not appear significant.
\newline\

```{r hist_plot23}

# Day of the Week

p <- ggplot(clean_data, aes(x=VIC_INC_DOW)) +
     geom_bar(fill="dark green") +
     labs(x="Day", y="Victim Records", title="Records By Day of Week") +
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

# Remove scientific notation
p + scale_y_continuous(labels = comma)
```
\newline\
The dip on Sunday is expected, but I was surprised by the lower count on Saturday.
\newline\

# 3. EDA - Look for correlation

## a) Bivariate Plots
```{r cor_plot}

ggplot(clean_data, aes(x=AGE_OF_VICTIM, y=CURRENT_POP1)) +
    geom_point(position="jitter") +
    labs(x="Age", y="Population", title="Victim Age vs Population") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
\newline\
Possible candidate for min-max scaling.
\newline\

## b) Test for correlation
```{r cor}
cor.test(clean_data$AGE_OF_VICTIM, clean_data$CURRENT_POP1, method="pearson")
```
\newline\
This numeric value indicates there is no correlation between victim's age and population of reporting district.
\newline\

# 4. Get Location Data & Export Final Dataset

## a) Import the Data
```{r geo_load, message=FALSE}

# Load geodetic city data
geo_data <- read_csv("Data/Location_Data/uscities.csv")

#head(geo_data)
```

## b) Re-Label fields

```{r new_geo_vars, include=TRUE}

# Rename fields so they can be joined
names(clean_data)[names(clean_data) == "CITY_NAME"] <- "CITY"
names(clean_data)[names(clean_data) == "STATE_ABBR"] <- "STATE" 
names(geo_data)[names(geo_data) == "city"] <- "CITY"
names(geo_data)[names(geo_data) == "state_id"] <- "STATE" 

# Convert city names to uppercase for matching
geo_data$CITY <- toupper(geo_data$CITY)

```

## c) Apply filters
```{r st_filter, include=TRUE}

# Limit only to high-reported states
# Filter both datasets to save memory
fil_clean_df <- filter(clean_data, STATE == "TN" | STATE == "MI" | STATE == "SC" |
                   STATE == "MA" | STATE == "OH" | STATE == "WA")

fil_geo_df <- filter(geo_data, STATE == "TN" | STATE == "MI" | STATE == "SC" |
                   STATE == "MA" | STATE == "OH" | STATE == "WA")

#head(fil_clean_df)
#head(fil_geo_df)
```

## d) Join Datasets
```{r geo_join, message=FALSE}

# Remove dataframes no longer needed to conserve memory
#rm("bat_data", "vic_data", "vic_new_data", "geo_data", "comb_df", "clean_data", "vic_date_data", "just_hom_df",
#   "off1_df", "off2_df", "off3_df", "off4_df", "off5_df", "off6_df", "off7_df",
#   "inj_df","inj1_df", "inj2_df", "inj3_df", "inj4_df", "inj5_df",
#   "asst_df","asst1_df", "asst2_df")

# Join cleaned and filtered data with city geodetic reference data
joint_df <- left_join(fil_clean_df, unique(fil_geo_df), c("CITY", "STATE"))

#head(joint_df)
#summary(joint_df)
```
   
## e) Clean and Output Data Frame
```{r clean_up, include=TRUE}

#rm("fil_clean_df", "fil_geo_df")

# Output clean and filtered file to save memory
# Can re-load for future use
write.csv(joint_df,"Data/crime_top6_states.csv")
```

