---
title: "DSC630 Final Project - Crime Analysis - Part2"
author: "Amie Davis"
date: "28 June, 2020"
output: 
  word_document: default
---
# Data Sources:
Uniform Crime Reporting Program Data: National Incident-Based Reporting System, [United States], 2016;
United States Federal Bureau of Investigation;
Inter-university Consortium for Political and Social Research (ICPSR), University of Michigan;
https://www.icpsr.umich.edu/icpsrweb/NACJD/NIBRS/
\newline\
Geodetic Data for US Cities:
https://simplemaps.com/data/us-cities

# References:
https://www.latlong.net
D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2. The R Journal, 5(1), 144-161. URL http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf


# Load Libraries

```{r libs, message=FALSE}
library(readr)
library(dplyr)
library(ggmap)
```

# 1. Prepare Data

## a) Import the Data
```{r load, message=FALSE}

# Load Cleaned data from Part1
crime_data <- read_csv("Data/crime_top6_states.csv")

#str(crime_data)
```

## b) Remove Unnecessary Columns

```{r drop2, include=TRUE}

# The following fields are not needed for visualizations and will be removed.

#  crime_data[ ,c(
#     "X1",
#     "VIC_INC_DATE",
#     "VICTIM_TYPE",
#     "ASSG_TYPE_OFFC",
#     "AGE_OF_VICTIM",
#     "SEX_OF_VICTIM",
#     "RACE_OF_VICTIM",
#     "ETHNIC_OF_VIC",
#     
#     "INJURY_TYPE1",
#     "INJURY_TYPE2",
#     "INJURY_TYPE3",
#     "INJURY_TYPE4",
#     "INJURY_TYPE5",
#  
#     "VIC_RESIDENT",
#     "NUM_RECS_PER_VICTIM",
#     "VIC_INC_YEAR",
#     "VIC_INC_MONTH",
#     "VIC_INC_DAY",
#     "VIC_INC_DOW",
#     "NUM_STATE_CODE",
#     "POP_GROUP",
#     "CTRY_DIVISION",
#     "CTRY_REGION",
#     "AGENCY_IND",
#     "CORE_CITY",
#     "FBI_OFFICE",
#     "JUDICIAL_DIST",
#     "CURRENT_POP1",
#     "UCR_COUNTY_CD1",
#     "MSA_CD1",
#     "LAST_POP1",
#     "FIPS_COUNTY1",
#     "Zip",
#     
#     "city_ascii",
#     "state_name",
#     "county_fips",
#     "county_name",
#     "county_fips_all",
#     "county_name_all",
#     "population",
#     "density",
#     "source",
#     "military",
#     "incorporated",
#     "timezone",
#     "ranking",
#     "zips"
#)] <- list(NULL)
  
#head(crime_data)
```

## c) Separate Offense Codes

```{r offense, include=TRUE}

# Create new df to include all offenses and their records
off1_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$OFF_CODE01))
off2_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$OFF_CODE02))
off3_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$OFF_CODE03))
off4_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$OFF_CODE04))
off5_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$OFF_CODE05))
off6_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$OFF_CODE06))
off7_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$OFF_CODE07))

names(off1_df)[names(off1_df) == "crime_data.OFF_CODE01"] <- "OFF_CODE"
names(off2_df)[names(off2_df) == "crime_data.OFF_CODE02"] <- "OFF_CODE"
names(off3_df)[names(off3_df) == "crime_data.OFF_CODE03"] <- "OFF_CODE"
names(off4_df)[names(off4_df) == "crime_data.OFF_CODE04"] <- "OFF_CODE"
names(off5_df)[names(off5_df) == "crime_data.OFF_CODE05"] <- "OFF_CODE"
names(off6_df)[names(off6_df) == "crime_data.OFF_CODE06"] <- "OFF_CODE"
names(off7_df)[names(off7_df) == "crime_data.OFF_CODE07"] <- "OFF_CODE"

off_df <- rbind(off1_df, off2_df, off3_df, off4_df, off5_df, off6_df, off7_df) 

head(off_df)
```

## d) Separate Aggrevated Assault Codes

```{r aggr, include=TRUE}

# Convert to cahracter
crime_data$ASSAULT_CIRC1 <- as.character(crime_data$ASSAULT_CIRC1)
crime_data$ASSAULT_CIRC2 <- as.character(crime_data$ASSAULT_CIRC2)

# Create new df to include all Aggravated Assault/Homicide Circumstances and their records
asst1_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$ASSAULT_CIRC1))
asst2_df <- na.omit(data.frame(crime_data$ORI, crime_data$INC_NUM, crime_data$ASSAULT_CIRC2))

names(asst1_df)[names(asst1_df) == "crime_data.ASSAULT_CIRC1"] <- "ASSAULT_CIRC"
names(asst2_df)[names(asst2_df) == "crime_data.ASSAULT_CIRC2"] <- "ASSAULT_CIRC"

asst_df <- rbind(asst1_df, asst2_df) 

head(asst_df)

```

## e) Join Datasets

```{r df_join, include=TRUE}

# Remove prefix from keys to match
names(off_df)[names(off_df) == "crime_data.ORI"] <- "ORI"
names(off_df)[names(off_df) == "crime_data.INC_NUM"] <- "INC_NUM"
#names(asst_df)[names(asst_df) == "crime_data.ORI"] <- "ORI"
#names(asst_df)[names(asst_df) == "crime_data.INC_NUM"] <- "INC_NUM"

# Join crime data with separate record for each offense
joint_off_df <- right_join(crime_data, off_df, c("ORI", "INC_NUM"))
#joint_all_df <- left_join(joint_off_df, asst_df, c("ORI", "INC_NUM"))

# Remove extra offense columns
  joint_off_df[ ,c(
     "OFF_CODE01",
     "OFF_CODE02",
     "OFF_CODE03",
     "OFF_CODE04",
     "OFF_CODE05",
     "OFF_CODE06",
     "OFF_CODE07"
)] <- list(NULL)

head(joint_off_df)
#summary(joint_off_df)

# Output joined file for use in modeling
write.csv(joint_off_df,"Data/crime_offenses.csv")

```

## f) Apply filters

```{r jh_filter, include=TRUE}

# Create subset to include only records marked as justifiable homicide
just_hom_df <- filter(joint_off_df, OFF_CODE == "09C")

summary(just_hom_df)
# There are 101 incidents of justifiable homicide reported to UCR in 2016 for the top 6 states
```

```{r agg_filter, include=TRUE}

# Create subset to include only records marked as aggravated assault
agg_asst_df <- filter(joint_off_df, OFF_CODE == "13A")

summary(agg_asst_df)
#head(agg_asst_df)
# There are 223,507 incidents of aggravated assaults reported to UCR in 2016 for the top 6 states
```

```{r act_filter, include=TRUE}

# Create subset to include only records marked as investigation of suspicious activity
act_type_df <- filter(joint_off_df, ACT_TYPE_OFFC == "7")

summary(act_type_df)
#head(act_type_df)
# There are 1,522 incidents of investigation of suspicious activity reported to UCR in 2016 for the top 6 states
```

```{r thresh, include=TRUE}

# Determine threshold for specific offense codes

# Justifiable Homicide by State
jh_state_df <- just_hom_df %>% group_by(STATE) %>% tally()
jh_state_df

# Aggrevated Assaults by State
aa_state_df <- agg_asst_df %>% group_by(STATE) %>% tally()
aa_state_df

# Total Offenses Reported by State
tot_state_df <- joint_off_df %>% group_by(STATE) %>% tally()
tot_state_df


# Justifiable Homicide by Population Group
jh_pop_df <- just_hom_df %>% group_by(POP_GROUP) %>% tally()
jh_pop_df

# Aggrevated Assaults by Population Group
aa_pop_df <- agg_asst_df %>% group_by(POP_GROUP) %>% tally()
aa_pop_df

# Total Offenses Reported by Population Group
tot_pop_df <- joint_off_df %>% group_by(POP_GROUP) %>% tally()
tot_pop_df

head(just_hom_df)
```

# 2. Visualizations for Features of Concern

\newline\
Using ggmap to use Google maps to display geodetic information.  The API key is passed, but is hidden from Markdown.
\newline

```{r goog, include=FALSE}
ggmap::register_google(key="")
```

## a) Tennessee

```{r tn_plots, message=FALSE}

# Create TN subset of all offense
tn_df <- filter(joint_off_df, STATE == "TN")

# Use Tennesee coordinates as center
tn_map <- ggmap(get_googlemap(
                  center = c(lon = -86.660156, lat = 35.860119),
                  zoom = 6, scale = 2,
                  maptype = "terrain",
                  color="color"))

# Build map
tn_map + 
  
  # Re-scale to focus on TN rectangle
  scale_y_continuous(limits=c(34.5, 36.5)) +
  scale_x_continuous(limits=c(-90.5, -82)) +

  # Add data points to the map
  geom_point(data = tn_df, 
             aes(x=lng, y=lat, color = OFF_CODE)) +
  labs(x="Longitude", y="Latitude", title="Incidents By Offense") +
  theme(legend.title = element_blank())


# Create TN subset of Justifiable Homicides
tn_spec_df <- filter(joint_off_df, STATE == "TN" & (OFF_CODE == "09C"))

# Build map
tn_map + 
  
  # Re-scale to focus on TN rectangle
  scale_y_continuous(limits=c(34.5, 36.5)) +
  scale_x_continuous(limits=c(-90.5, -82)) +

  # Add data points to the map
  geom_point(data = tn_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="red",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Justifiable Homicides")


# Create TN subset of Suspicious Activity Reporting
tn_spec_df <- filter(joint_off_df, STATE == "TN" & (ACT_TYPE_OFFC == "7"))

# Build map
tn_map + 
  
  # Re-scale to focus on TN rectangle
  scale_y_continuous(limits=c(34.5, 36.5)) +
  scale_x_continuous(limits=c(-90.5, -82)) +

  # Add data points to the map
  geom_point(data = tn_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="yellow",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Suspicious Activity Reporting")
```

## b) Michigan

```{r mi_plots, message=FALSE}

# Create MI subset of all offense
mi_df <- filter(joint_off_df, STATE == "MI")

# Use Michigan coordinates as center
mi_map <- ggmap(get_googlemap(
                  center = c(lon = -84.506836, lat = 44.182205),
                  zoom = 6, scale = 2,
                  maptype = "terrain",
                  color="color"))

# Build map
mi_map + 
  
  # Re-scale to focus on MI rectangle
  scale_y_continuous(limits=c(41.5, 47.5)) +
  scale_x_continuous(limits=c(-90.5, -82)) +

  # Add data points to the map
  geom_point(data = mi_df, 
             aes(x=lng, y=lat, color = OFF_CODE)) +
  labs(x="Longitude", y="Latitude", title="Incidents By Offense") +
  theme(legend.title = element_blank())


# Create MI subset of Justifiable Homicides
mi_spec_df <- filter(joint_off_df, STATE == "MI" & (OFF_CODE == "09C"))

# Build map
mi_map + 
  
  # Re-scale to focus on MI rectangle
  scale_y_continuous(limits=c(41.5, 47.5)) +
  scale_x_continuous(limits=c(-90.5, -82)) +

  # Add data points to the map
  geom_point(data = mi_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="red",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Justifiable Homicides")


# Create MI subset of Suspicious Activity Reporting
mi_spec_df <- filter(joint_off_df, STATE == "MI" & (ACT_TYPE_OFFC == "7"))

# Build map
mi_map + 
  
  # Re-scale to focus on MI rectangle
  scale_y_continuous(limits=c(41.5, 47.5)) +
  scale_x_continuous(limits=c(-90.5, -82)) +

  # Add data points to the map
  geom_point(data = mi_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="yellow",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Suspicious Activity Reporting")
```

## c) South Carolina

```{r sc_plots, message=FALSE}

# Create SC subset of all offense
sc_df <- filter(joint_off_df, STATE == "SC")

# Use South Carolina coordinates as center
sc_map <- ggmap(get_googlemap(
                  center = c(lon = -81.163727, lat = 33.836082),
                  zoom = 6, scale = 2,
                  maptype = "terrain",
                  color="color"))

# Build map
sc_map + 
  
  # Re-scale to focus on SC rectangle
  scale_y_continuous(limits=c(32, 35.5)) +
  scale_x_continuous(limits=c(-83.5, -78)) +

  # Add data points to the map
  geom_point(data = sc_df, 
             aes(x=lng, y=lat, color = OFF_CODE)) +
  labs(x="Longitude", y="Latitude", title="Incidents By Offense") +
  theme(legend.title = element_blank())


# Create SC subset of Justifiable Homicides
sc_spec_df <- filter(joint_off_df, STATE == "SC" & (OFF_CODE == "09C"))

# Build map
sc_map + 
  
  # Re-scale to focus on SC rectangle
  scale_y_continuous(limits=c(32, 35.5)) +
  scale_x_continuous(limits=c(-83.5, -78)) +

  # Add data points to the map
  geom_point(data = sc_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="red",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Justifiable Homicides")


# Create SC subset of Suspicious Activity Reporting
sc_spec_df <- filter(joint_off_df, STATE == "SC" & (ACT_TYPE_OFFC == "7"))

# Build map
sc_map + 
  
  # Re-scale to focus on SC rectangle
  scale_y_continuous(limits=c(32, 35.5)) +
  scale_x_continuous(limits=c(-83.5, -78)) +

  # Add data points to the map
  geom_point(data = sc_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="yellow",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Suspicious Activity Reporting")
```

## d) Massachusetts

```{r ma_plots, message=FALSE}

# Create MA subset of all offense
ma_df <- filter(joint_off_df, STATE == "MA")

# Use Massachusetts coordinates as center
ma_map <- ggmap(get_googlemap(
                  center = c(lon = -71.382439, lat = 42.407211),
                  zoom = 6, scale = 2,
                  maptype = "terrain",
                  color="color"))

# Build map
ma_map + 
  
  # Re-scale to focus on MA rectangle
  scale_y_continuous(limits=c(41.5, 43)) +
  scale_x_continuous(limits=c(-74, -70)) +

  # Add data points to the map
  geom_point(data = ma_df, 
             aes(x=lng, y=lat, color = OFF_CODE)) +
  labs(x="Longitude", y="Latitude", title="Incidents By Offense") +
  theme(legend.title = element_blank())


# Create MA subset of Justifiable Homicides
ma_spec_df <- filter(joint_off_df, STATE == "MA" & (OFF_CODE == "09C"))
#summary(ma_spec_df)
# There are were no Justifiable Homicides reported in MA for 2016.


# Create MA subset of Suspicious Activity Reporting
ma_spec_df <- filter(joint_off_df, STATE == "MA" & (ACT_TYPE_OFFC == "7"))

# Build map
ma_map + 
  
  # Re-scale to focus on SC rectangle
  scale_y_continuous(limits=c(41.5, 43)) +
  scale_x_continuous(limits=c(-74, -70)) +

  # Add data points to the map
  geom_point(data = ma_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="yellow",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Suspicious Activity Reporting")
```

## e) Ohio

```{r oh_plots, message=FALSE}

# Create OH subset of all offense
oh_df <- filter(joint_off_df, STATE == "OH")

# Use Ohio Carolina coordinates as center
oh_map <- ggmap(get_googlemap(
                  center = c(lon = -82.996216, lat = 40.367474),
                  zoom = 6, scale = 2,
                  maptype = "terrain",
                  color="color"))

# Build map
oh_map + 
  
  # Re-scale to focus on OH rectangle
  scale_y_continuous(limits=c(38, 42)) +
  scale_x_continuous(limits=c(-85, -80)) +

  # Add data points to the map
  geom_point(data = oh_df, 
             aes(x=lng, y=lat, color = OFF_CODE)) +
  labs(x="Longitude", y="Latitude", title="Incidents By Offense") +
  theme(legend.title = element_blank())


# Create OH subset of Justifiable Homicides
oh_spec_df <- filter(joint_off_df, STATE == "OH" & (OFF_CODE == "09C"))
#summary(oh_spec_df)
# There are were no Justifiable Homicides reported in OH for 2016.


# Create OH subset of Suspicious Activity Reporting
oh_spec_df <- filter(joint_off_df, STATE == "OH" & (ACT_TYPE_OFFC == "7"))

# Build map
oh_map + 
  
  # Re-scale to focus on SC rectangle
  scale_y_continuous(limits=c(38, 42)) +
  scale_x_continuous(limits=c(-85, -80)) +

  # Add data points to the map
  geom_point(data = oh_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="yellow",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Suspicious Activity Reporting")
```

## f) Washington State

```{r wa_plots, message=FALSE}

# Create WA subset of all offense
wa_df <- filter(joint_off_df, STATE == "WA")

# Use Washington coordinates as center
wa_map <- ggmap(get_googlemap(
                  center = c(lon = -120.740135, lat = 47.751076),
                  zoom = 6, scale = 2,
                  maptype = "terrain",
                  color="color"))

# Build map
wa_map + 
  
  # Re-scale to focus on SC rectangle
  scale_y_continuous(limits=c(45.5, 49)) +
  scale_x_continuous(limits=c(-124.5, -116.5)) +

  # Add data points to the map
  geom_point(data = wa_df, 
             aes(x=lng, y=lat, color = OFF_CODE)) +
  labs(x="Longitude", y="Latitude", title="Incidents By Offense") +
  theme(legend.title = element_blank())


# Create WA subset of Justifiable Homicides
wa_spec_df <- filter(joint_off_df, STATE == "WA" & (OFF_CODE == "09C"))

# Build map
wa_map + 
  
  # Re-scale to focus on WA rectangle
  scale_y_continuous(limits=c(45.5, 49)) +
  scale_x_continuous(limits=c(-124.5, -116.5)) +

  # Add data points to the map
  geom_point(data = wa_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="red",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Justifiable Homicides")


# Create WA subset of Suspicious Activity Reporting
wa_spec_df <- filter(joint_off_df, STATE == "WA" & (ACT_TYPE_OFFC == "7"))

# Build map
wa_map + 
  
  # Re-scale to focus on WA rectangle
  scale_y_continuous(limits=c(45.5, 49)) +
  scale_x_continuous(limits=c(-124.5, -116.5)) +

  # Add data points to the map
  geom_point(data = wa_spec_df, 
             aes(x=lng, y=lat, size=3),
             color="yellow",
             show.legend = FALSE) +
  labs(x="Longitude", y="Latitude", title="Suspicious Activity Reporting")
```
